Visualizing Functional Data

### 2017-05-09 paper/sections/section0d.tex
sections are awesome

\subsection{Multivariate Observations}
If the variables are independent, then small multiples (insert citation) or
visualizations for a smaller number of variables can be used to visualize
them. But this technique is limited to a somewhat small number of variables
(maybe 20 (find paper)) before the number of multiples is overwhelming and
impedes decision making. Instead, there are some techniques to look at a large
set of independent variables at once, such as ....

%% large scale box: letter, spark

>>>>

\subsection{Multivariate Observations}
If the variables are independent, then small multiples (insert citation) or
visualizations for a smaller number of variables can be used to visualize
them. But this technique is limited to a somewhat small number of variables
(maybe 20 (find paper)) before the number of multiples is overwhelming and
impedes decision making. Instead, there are some techniques to look at a larger
set of independent variables at once, such as ....

The simiplest of these is the bubble plot (insert citation), which is a simple
extension of the scatter plot where the size of the scatter point varies
relative to some ordinal variable. Varying the marker style and colors adds two
more levels of (insert citation on cramming in all the viz on scatters), and
the scatter can be 3d, but the maximum number of variables tops out at about 6
and the graph has a lot of visual noise.

Instead a typical approach is to visualize the piecewise relationships. The
simpliest of these are a matrix of scatterplots (insert scatter matrix
citation), but the relationships can get drowned out when there are a large
number of observations. So instead, a common approach is to use parralel
coordinates (insert citation) wherein each variable is an axis and lines are
drawn between the value of variable 1 and the value of variable 2. Bundles of
lines indicate very common trends between the two variables, and many
implementations of parallel coordianets have movable axis so users can explore
each pairwise relationship (insert citation). Radial graphs are a type of 
parallel coordinates where all the variables are linked to each other; this
forms a circular type graph wherin the area of the graph is used as an
indicator of the relationship amongst these variables (insert citation). There
are a familiy of these sorts of radial visualizations...

%%show picture parallel coordinate

There are more streamlined approaches
such as  %% radial, parallel coordinates, factor plot, but

Even the best of these visualization techniques are limited to about 5 or 6 variables
%% large scale box: letter, spark

### 2017-05-09 paper/sections/section1d.tex
sections are awesome

\begin{figure}
  \includegraphics[width=.75\linewidth]{contour_weather}
  \caption{\label{fig:countour}}
\end{figure}
While functional bag, box, and HDR plots all show outliers, each method is most
successful visualizing a specific type of outlier. Countour
boxplots\cite{whitaker2013} aim instead to capture the variablity of the
uncertainity via a band depth ordering metric. Each ensemble members band depth
is computed as sum of the probabilities that the observations in any given
ensemble fall within the max-min envelope defined by any two other
ensembles. The bands are then sorted by band depth such that the median is the
ensemble with about 50\% of its members withen all envelopes formed by other
bands (so most centered). Outer bands are chosen according to the task at
hand. In \cite{whitaker2013}, they apply the countour boxplots to temperature
visualizations. The authors argue that countour boxplots are an improved visual
idiom over the traditional spagehtti plots (%insert citation for them...)
because the spagehetti plots get nosy as the number of plots increase
and so it's hard to tease out specific patterns. Instead in
Figure~\ref{fig:countour}, the aut\
hors
remove most of the bands and instead visualize an outlier envelope(light gray)
and a more central envelope(dark gray). It retains much of the information of
the spagethhit, but removes the visual noise of the lines.

@@

\subsection{Multivariate Observations}
LDA, CCA, 
\subsection{Relationships Between Variables}

\subsubsection{Bivariate}
surface map, functional correlations, planes?
\subsection{Multivariate Observations}
PLS, functional PCA

>>>>

\begin{figure}
  \includegraphics[width=.75\linewidth]{contour_weather}
  \caption{\label{fig:countour}}
\end{figure}
While functional bag, box, and HDR plots all show outliers, each method is most
successful visualizing a specific type of outlier. Countour
boxplots\cite{whitaker2013} aim instead to capture the variablity of the
uncertainity via a band depth ordering metric. Each ensemble members band depth
is computed as sum of the probabilities that the observations in any given
ensemble fall within the max-min envelope defined by any two other
ensembles. The bands are then sorted by band depth such that the median is the
ensemble with about 50\% of its members withen all envelopes formed by other
bands (so most centered). Outer bands are chosen according to the task at
hand. In \cite{whitaker2013}, they apply the countour boxplots to temperature
visualizations. The authors argue that countour boxplots are an improved visual
idiom over the traditional spagehtti plots (%insert citation for them...)
because the spagehetti plots get nosy as the number of plots increase
and so it's hard to tease out specific patterns. Instead in
Figure~\ref{fig:countour}, the authorsremove most of the bands and instead visualize an outlier envelope(light gray)
and a more central envelope(dark gray). It retains much of the information of
the spagethhit, but removes the visual noise of the lines.

@@

surface map, functional correlations, planes?

\subsection{Multivariate Observations}
LDA, CCA, 
PLS, functional PCA
